---
title: "Top 250 Football Transfers"
output:
  html_document:
    df_print: paged
---

<br />
<br />
<br />
**LIBRARIES**

```{r}
library(tidyverse,warn.conflicts = FALSE)
library(ggplot2,warn.conflicts = FALSE)
library(gganimate,warn.conflicts = FALSE)
library(dplyr,warn.conflicts = FALSE)
library("readxl")
```
<br />
<br />


**PATH for the Excel file**
```{r}
TransferFees_Football<-read_excel("C:\\Users\\Adrian\\Desktop\\DAB_Datasets\\TransferFees_Football.xlsx")
```
<br />
<br />



**Data Cleaning**
<br />
In the column League_to , we have 1.Bundesliga- First Div , 2.Bundesliga-Second Div etc. <br />
Since I am only making use of the First Div football,I have cleaned the variable 
League_to from "1.Bundesliga" to "Bundesliga" <br />
```{r}
TransferFees_Football<-TransferFees_Football%>%
  mutate(League_to = replace(League_to, League_to == "1.Bundesliga", "Bundesliga"))

#check if 1.Bundesliga  is replaced
TransferFees_Football%>%
  filter(League_to=="1.Bundesliga")
```



**Dataset**

```{r}
glimpse(TransferFees_Football)
```
The dataset consists of 4700 observations and 9 variables.It gives us information about the top 250 Football Transfers between 2000-2018  <br />
The variables are - <br />
1)Name           &nbsp; (chr) &nbsp;   - Name of the player which was sold <br />
2)Position       &nbsp; (chr) &nbsp;   - Playing position of player <br />
3)Player Age     &nbsp; (dbl) &nbsp;   - Age of the player <br />
4)Team_from      &nbsp; (chr) &nbsp;   - The selling team <br />
5)League_from    &nbsp; (chr) &nbsp;   - The selling League <br />
6)Team_to        &nbsp; (chr) &nbsp;   - The buying team <br />
7)League_to      &nbsp; (chr) &nbsp;   - The buying League <br />
8)Season         &nbsp; (chr) &nbsp;   - Season of the transfer <br />
9)Transfer_fee   &nbsp; (dbl) &nbsp;   - The selling price of the player<br />
<br />
<br />
<br />







**Provide output for the head(), tail(), and summary() commands for your data set.**


**head()**
<br />
The head() function in R is used to display the first n values of the dataset. We can specify the number of observations we want to view using custom head
Example - [head(TransferFees_Football),n]
In this example we can use any integer value for n. The number of observations to be returned depends on the value of n

In our case we have returned the top 6 (default-6) observations of our data set.  
```{r}
head(TransferFees_Football)
```

<br />
<br />
<br />



**tail()**
<br />
The tail() function in R is used to display the last n values of the dataset. We can specify the number of observations we want to view using custom tail
Example - [tail(TransferFees_Football),n]
In this example we can use any integer value for n. The number of observations to be returned depends on the value of n

In our case we have returned the bottom 6 (default-6) observations of our data set.  

```{r}
tail(TransferFees_Football)
```

<br />
<br />
<br />

**summary()**
<br />
It is used to summarize all the values in  the dataset.It provides a summary with the objective of a quick glance.
The summary below provides mean,median,min,max,Data Type,length etc.
```{r}
summary(TransferFees_Football)
```

<br />
<br />
<br />

--------------------------------------------------------------------------------

**Create 2 exploratory type questions and provide the code and output for the corresponding answers. Each answer must use at least two of the following dplyr verbs: filter(), mutate(), summarize(), and group_by(). (Along the lines of Exercises 11-16 in R4DS_Chapter_5d.html.) Please avoid overly simplistic questions.**


<br />
<br />
**1.Among the top 5 leagues in Europe, which leagues have payed the            highest transfer fees and  what is the avg age of those                   players**

**Define the top 5 Leagues in Europe**
<br />
The top 5 leagues in Europe constists of  <br />
Serie A - First Division of Italy   <br />
LaLiga - First Division of Spain <br />
Premier League  - First Division of England   <br />
Ligue 1 - First Division of France  <br />
1.Bundesliga - First Division of  Germany <br />
```{r}
Top_Euro_Countries<-c("Serie A","LaLiga",'Premier League','Ligue 1','Bundesliga')
```

***Extract the top 5 Leagues in Europe with their Highest transfer fees and average age ***
```{r}
   Q1 <- TransferFees_Football %>% 
          filter(League_to %in% Top_Euro_Countries) %>% 
            group_by(League_to) %>% 
          summarise(Mean_Age=(mean(`Player Age`)),Mean_Transfer=mean((Transfer_fee))) %>% 
          mutate(Round_Age=round(Mean_Age)) %>%
          arrange(desc(Mean_Age),desc(Mean_Transfer)) 

Q1
```
**Description**
  In the above question we need to find max transfer and the average age of players in the transfer market based on their respective league.<br />
  Verbs used - filter(),group_by(),summarize(),mutate()    <br />
  1)Filter the top 5 Leagues in Europe.  <br />
  2)Group By the league.     <br />
  3)Summarize the mean of Age and the transfer fee.<br />
  4)The Mutate function is used to add the rounded value of the Mean_Age  <br />
  5)To get highest fees- order by descending values.<br />

**Observation**
It can be clearly noted that the above mentioned leagues in Europe prefer players in
their mid 20s. The Premier League has spent the most in the market whereas the Bundesliga has invested the least among the top 5 leagues


---
---

</br>
</br>
</br>

**2.From the year 2000-2018 which clubs has spent  over 100 million in the transfer market and name the player**

```{r}

  sales_over100mil<-  TransferFees_Football %>%
        filter(Transfer_fee>100000000) %>% 
        group_by(Team_from,Transfer_fee,Team_to) %>% 
        summarise(Name=unique(Name)) %>% 
        arrange(desc(Transfer_fee))

sales_over100mil
  
```

**Description**
 In this question we need the top  teams which spent over 100 mil <br />
Verbs used - filter(),group_by(),summarize(),arrange()    <br />
1)First we need to filter the teams which have spent over 10 million.   <br />
2)Group by to get the Team that spent the amount(Team_to)<br />
3)Summarize to get the player name    <br />
4)Arrange in descending order to get the maximum first.    <br />

**Observation**
Based on the exploratory output.We can conclude that Paris SG,FC Barcelona,Juventus,
Man Utd & Real Madrid have spent the most in the transfer market. Another observation is that FC Barcelona,Juventus,Real Madrid have  received 100mil+ from player sales.


---
---


---------------------------------------------------------------------------------

